EuPPollNet: A European database of plant-pollinator networks

### 2024-05-22 Manuscript/Figures_and_Tables/Figure_6_Updated.Rmd
edits

---
title: "EuPPollNet: A European database of plant-pollinator networks"
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}

bibliography: references.bib
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(sf) #For handling coordinates and plotting
library(scales) #For plotting (decimals on axes)
library(lubridate) #To operate with dates
library(tidyr) #Reshape data (wide format)
library(stringr) #Fix/modify strings
library(forcats)
library(readr)
library(phytools)
#Load libraries for plotting
library(giscoR) #Maps
library(ggstar) #cool shapes as points
library(ggplot2) #general plotting
library(ggtree) #phylo trees
library(ggpattern) #cool shapes in barplots
library(patchwork) #plot organisation
library(cowplot) #plot organisation
library(viridis) #colours
library(ggnewscale)
library(ggtreeExtra)
library(kableExtra)

#In case of conflicts select the main dplyr functions
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

#Read data
data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
mutate(Network_id = paste0(Study_id, Network_id))  
#Bee-syrphid Masterlist
master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
#Load butterfly masterlist
master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
                    header = T)
#Prepare table with data structure (as it would be made available to others)
safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")

#Work around to get the three cols
#Note that month 1 and day 1 indicates the presence of a missing value
#At least in Europe :)
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
                Month = lubridate::month(Date), 
                Day = lubridate::day(Date))

#Country number
country_number = data %>% 
select(Country) %>% 
n_distinct()
#Study number
study_number = data %>% 
select(Study_id) %>% 
n_distinct()
#Average study number
studies_by_country = data %>% 
group_by(Country) %>% 
summarise(study_number = n_distinct(Study_id)) %>% 
summarise(mean(study_number)) %>% 
pull()
#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
#Sampling method
Sampling_method = data %>% 
group_by(Study_id) %>% 
distinct(Sampling_method) %>%  
ungroup() %>% 
count(Sampling_method) %>% 
mutate(percent_method = n/sum(n) *100) %>% 
mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
#1st most common sampling methodology
most_common_method = Sampling_method %>% 
filter(percent_method == max(percent_method)) 
#Minimum year
min_year = data %>% 
select(Year) %>% 
slice(which.min(Year)) %>% 
pull()
#Maximum year
max_year = data %>% 
select(Year) %>% 
slice(which.max(Year)) %>% 
pull()
#Number of interactions
interaction_number = nrow(data)
#Check number of accepted pollinators
pollinator_spp_number = data %>% 
select(Pollinator_rank, Pollinator_accepted_name)%>%
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name) %>% 
n_distinct()
#Check number of plant species
plant_spp_number = data %>% 
select(Plant_rank, Plant_status, 
Plant_matchtype, Plant_accepted_name) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name) %>% 
n_distinct()
#Extract percentage of studies conducted during a single year
dates = data %>% 
select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
group_by(Study_id) %>% 
summarise(Count = n_distinct(Year))
#Filter studies with only one year
one_year = dates %>% 
filter(Count == 1) %>% 
pull() %>% 
length() / nrow(dates) * 100
#Obtain date information
dates = data %>% 
select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))

dates1 = dates %>% 
group_by(Study_id, Network_id, Sampling_method) %>% 
summarise(Min_date = min(Date), 
          Max_date= max(Date), 
          Sampling_days = n_distinct(Date)) %>% 
mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
       .before=Study_id) %>% 
arrange(Study_number) %>% 
mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
TRUE ~ NA)) %>% 
mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
TRUE ~ NA))
#Extract min and max number of sampling days
min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
#Extract average of sampling days across networks and studies
average_sampling_days = dates1 %>% 
ungroup() %>% 
group_by(Study_id) %>% 
summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
ungroup() %>% 
summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
pull()
#Check number of distinct sampling sites
n_sites = data %>% 
select(Latitude, Longitude) %>% 
distinct() %>% 
nrow(.)
#Check number of distinct bioregions
n_bioregion = data %>% 
select(Bioregion) %>% 
distinct()  %>% 
nrow(.)
#Read author list 
authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
#Print authors
authors = authors %>% 
summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
pull()
#Read affiliation list 
aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
aff_list = aff_list %>% 
summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
pull()

#Interactions at species level
species_level = data %>% 
filter(Plant_rank == "SPECIES") %>% 
filter(Pollinator_rank == "SPECIES") %>% 
filter(Plant_status == "ACCEPTED") %>% 
filter(Pollinator_status == "ACCEPTED") %>% 
nrow(.)
species_level_int = species_level / nrow(data) * 100
#Read here plant and pollinator coverage
bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")

#Calculate mean coverage and deviation
coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
mean_coverage = mean(coverage)
sd_coverage = sd(coverage)

#Main orders percentage
o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
main_orders = data %>% 
filter(Pollinator_order %in% o)
main_orders_percent = nrow(main_orders) / nrow(data) * 100

#Prepare table with all info 
#Create variables
vars = colnames(safenet_database)
#Create table
d = tibble(Variable = vars)

d = d %>% 
mutate(Variable = str_replace(Variable, "Latitude", "Latitude-Longitude")) %>% 
filter(!Variable == "Longitude")

#Add descriptor
my_tibble = d %>%
  mutate(Description = case_when(
    Variable == "Study_id" ~ "Identifier of the study",
    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
    Variable == "Bioregion" ~ "European biogiographical regions",
    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
    Variable == "Latitude-Longitude" ~ "Coordinates of the observed interaction in decimal degrees",
    Variable == "Date" ~ "Year, month and day when the observation took place",
    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
    Variable == "Plant_old_name" ~ "Plant species name given by the authors",
    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
    Variable == "Pollinator_old_name" ~ "Pollinator species name given by the authors",
    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",

    TRUE ~ "Other"
  ))

#Check number of species and percentage of interactions of poll. groups
main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

order_info = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
          Species = n_distinct(Pollinator_accepted_name))

Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
select(Interactions) %>% pull()

#Check percentage of studies with flower counts
floral_counts_percent = data  %>%
select(Study_id, Flower_data) %>% 
distinct() %>% 
group_by(Flower_data) %>% 
summarise(Percentage = n()/ nrow(.)*100) %>% 
filter(Flower_data == "Yes")

#Check percentage of apis mellifera records
all_records = data %>% nrow(.) 
apis_records = data %>% 
filter(Pollinator_accepted_name == "Apis mellifera") %>% 
nrow(.)
apis_percent = apis_records/all_records * 100
#Check average pecentage of interactions of apis by net
a = data %>% 
group_by(Network_id) %>% 
summarise(n_interactions = length(Pollinator_accepted_name))
b = data %>% 
group_by(Network_id) %>% 
filter(Pollinator_accepted_name == "Apis mellifera") %>% 
summarise(n_interactions_apis = length(Pollinator_accepted_name))
c = left_join(a,b)
apis_average_network = c %>% 
mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
ungroup() %>% 
summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
pull()

\vspace*{-14mm}

`r authors`

**Corresponding author**= barragansljose@gmail.com\
*All authors excluding the first three and last are ordered alphabetically*

\begingroup
\small

`r aff_list`

\small
\endgroup

\pagebreak

# Abstract
\vspace*{-5mm}

**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. Indeed, the intricate network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning. However, rapid human-induced environmental changes are compromising its long-term persistence. One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time. Here, we present the EuPPollNet database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time. This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences, while also providing other ecological variables of interest. Furthermore, this work provides an in-depth assessment of the taxonomic and sampling coverage of the database at the European level, complemented by analyses of key structural properties in plant-pollinator networks. We hope this database can help researchers to: 1) identify taxonomic, ecological, and geographical gaps of knowledge on plant-pollinator interactions; and 2) explore the impacts of global change on plant-pollinator networks to guide future conservation planning for both plant and pollinator species.

**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_sites, big.mark=",")` distinct locations (i.e., plant-pollinator networks), which belong to `r study_number` different studies distributed across `r country_number` European countries. In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided.

**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions. All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).

**Time period and grain:** Species interaction data was recorded between `r min_year` and `r max_year`. All records are time-referenced and most of the studies documented interactions in a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).

**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species. The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphids species at the European level.

**Software format:** The database was built with R software and is stored as ".rds" and ".csv" formats. The construction of the database is fully reproducible and can be accessed at the following PERMANENT LINKS.

**KEYWORDS**\
\begingroup
\small
plant-pollinator networks, plant-pollinator interactions, flowering plants, Angiosperms, pollinators, nestedness, connectance
\endgroup
\pagebreak

# 1 | INTRODUCTION

The interaction between plants and pollinators is one of the most well-documented mutualisms on Earth. Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and the human economy. The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021]. Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species and guiding conservation planning [@heberling2021]. However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists along with occurrence data, is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding the fate of the species within an ecosystem [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, originating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate the state of the art of plant-pollinator interactions at European level, which will guide research efforts, conservation planning and will set a foundation for future global change research.

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties across them, such as truncated power-law degree distributions [@jordano1987] or nestedness [@bascompte2003]. In addition, this type of large scale analyses can help understanding landscape level processes that cannot be explored at the community level, such as ecological patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023]. Although macro-ecological approaches make significant contributions to knowledge, they tend to be rare, and are strongly influenced by the spatio-temporal availability and nature of the data [@burkle2011; @trojelsgaard2013]. For instance, plant-pollinator studies tend to differ in sampling effort and methodology which can impact the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016]. Further, despite little sampling effort could capture most of the relevant functional species [@hegland2010], most plant-pollinator networks have unobserved interactions because undersampling [@olesen2011; @chacoff2012]. Thus, the lack of strong spatio-temporal coverage, along with the current intrinsic limitations of sampling plant-pollinator networks, highlight the need of keep gathering and integrating informative species interaction data to properly unravel the different ecological processes that shape plant-pollinator interactions at large scales.

Here, we present the EuPPollNet database, an acronym derived from European plant-pollinator networks, which contains harmonized information on interaction data of plants and pollinators at European level. The primary focus of the pollinator taxonomic groups is on the main orders of entomofauna that visit and pollinate flowering plants in Europe. These include insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera [@potts2015], which comprise almost the totality of recorded interactions in the EuPPollNet database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). To understand the scope of the database, we examine the taxonomic and sampling coverage of the different plant and pollinator species at European level with the help of the most up-to-date species checklists and rarefaction analyses. In addition, we investigate how two key structural metrics of plant-pollinator networks, such as connectance and nestedness, change across the latitudinal range of studies and bioclimatic regions. Finally, we also examine whether plant-pollinator networks fulfill the expectation of being more nested than null expectations. EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that allow better control for the ecological context and sampling methods. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows it to be reused and keep building on it over time. This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions. We expect that this database will help to evaluate macro-ecological processes and current gaps of plant-pollinator interactions at European level.

# 2 | METHODS

## Data acquisition

This database is the result of one of the working packages of the European project Safeguard. The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent. Data acquisition followed a non-systematic approach. First, data was directly asked to members of the Safeguard project and then, the request was extended to data owners outside of the project. These other studies were identified by direct communication with other colleagues and by directly searching studies on Google Scholar of under-represented regions within the database. To keep high quality standards that will allow robust future ecological research, we only included studies meeting the following criteria: 1) studies that contained time- and geo-referenced records of plant-pollinator interactions; and 2) studies with phyto-centric plant-pollinator networks with quantitative visitation data.

## Dataset description

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` on `r country_number` different countries (**Figure 1a** and **Figure 1b**). The different studies differ in sampling effort and methodology, although most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and documented interactions mostly by using `r  most_common_method[[1]]` as sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions, considering interaction as the contact of a given pollinator to the reproductive structure of a particular plant. Most of the pollinator species in the database belong to the orders Hymenoptera and Diptera, each comprising approximately 1,000 species. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably,  _Apis mellifera_ represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "\\textbf{Figure 1}. (\\textbf{a}) Approximate locations of the studies of the EuPPollNet database showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study. The sizes of these shapes are proportional to the respective species counts. For visualization purposes, we have focused only on the European region with studies and selected a single location per study. (\\textbf{b}) Number of studies by year within the database. (\\textbf{c}) Number of species and interactions across the 4 main pollinator orders at European level."}

#Fix dates
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
                Month = lubridate::month(Date), 
                Day = lubridate::day(Date))

#Prepare coordinates
for_plotting = data %>% 
select(Study_id, Latitude, Longitude) %>% 
distinct()
#Get countries and transfrom coordinates
countries <- gisco_get_countries(
resolution = 20) %>%
st_transform(3035) %>% 
sf::st_as_sf(crs = 4326)

#Number of plant and poll species per network
#-----------------------#
#Pollinators------
#-----------------------#
#Check unique number of pollinator species
PollNumber_by_ID = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
#Merge now with spp counts by network dataset
poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
poll_data$coords <- data.frame(
Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
#Set coordinates
poll_data$coords_sf = st_as_sf(poll_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
poll_data1 =  poll_data %>% 
distinct(Study_id, .keep_all = TRUE)

#-----------------------#
#Plants-----
#-----------------------#
#Check unique number of plant species
PlantNumber_by_ID = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Plant_counts = n_distinct(Plant_accepted_name))
#Merge now with spp counts by network dataset
plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
plant_data$coords <- data.frame(
Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
#Set coordinates
plant_data$coords_sf = st_as_sf(plant_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
plant_data1 =  plant_data %>% 
distinct(Study_id, .keep_all = TRUE)

#Add vars
plant_data1$Group = "Plant"
poll_data1$Group = "Pollinator"
#Rename cols
plant_data2 = plant_data1 %>% 
rename(Counts = Plant_counts)
poll_data2 = poll_data1 %>% 
rename(Counts = Pollinator_counts)
#Bind rows
all_data = bind_rows(plant_data2, poll_data2)

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
           size=Counts, starshape = Group, color= Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("top"),
      legend.box = "vertical",
      legend.background = element_blank(),
      legend.box.background = element_blank(),
      legend.key = element_blank(), 
      legend.title = element_text(face ="bold",  size = 6),
      legend.box.spacing = unit(0.2, "cm"),
      legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
      axis.text = element_text(size=6,                  face="bold"),
      axis.title = element_text(face="bold", size = 8),
         legend.text = element_text(size = 6),                           legend.direction="horizontal",
         legend.spacing.y = unit(-0.3, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
scale_color_manual(values=c("#009E73", "#D55E00"),
         guide="none") +
scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
             guide=guide_legend(reverse=TRUE,
             override.aes=list(
                 size=2,
                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
  ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
ggtitle("(a)") +
theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))

#----------------#
#Pollinators
#----------------#
#First select cols of rank, accepted name and study ID
#Then select unique levels by study ID
#Now sum levels (maximum can be as maximum number of studies)
poll_spread = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
distinct(Pollinator_accepted_name) %>% 
ungroup() %>% 
count(Pollinator_accepted_name) %>% 
rename(n_studies_shared = n) %>% 
arrange(-n_studies_shared) %>% 
mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))

#Seelect quantile 50% from the total number of species
spp_number = poll_spread %>%  
select(Pollinator_accepted_name) %>% 
n_distinct()
df = seq(1, spp_number)
quantile_50 = quantile(df, 0.5)

#Obtain most common pollinator for plotting
most_common_poll = poll_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
select(Pollinator_accepted_name) %>% pull()

#Plot
#p1 = ggplot(poll_spread, 
#aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
#geom_bar(stat = "identity", 
#           color = "#D55E00",
#           lwd = 0.015, fill = "#D55E00") +
#ylab("Shared pollinators (%)") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#  axis.ticks.x=element_blank(),
#  axis.ticks.y = element_line(size = 0.2),
#  plot.title = element_text(size = 8, face = "bold"), 
#  axis.text.y = element_text(size=5),
#  panel.grid.major.x = element_blank(),
#  panel.grid.minor.y = element_blank(),
#  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
#xlab(NULL) +
#ylab(NULL) + 
#ggtitle("Pollinators")+
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#color = "black", size=0.25) + 
#annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = #1.5) +
#annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  #fontface = 'italic', size = 1.5)+
#scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
#annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
#coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, #max(poll_spread$Percent_total)), clip="off")
#
##----------------#
##Plants
##----------------#
#plant_spread = data %>% 
#select(Plant_rank, Plant_accepted_name, Study_id) %>% 
#group_by(Study_id) %>% 
#filter(Plant_rank == "SPECIES") %>% 
#distinct(Plant_accepted_name) %>% 
#ungroup() %>% 
#count(Plant_accepted_name) %>% 
#rename(n_studies_shared = n) %>% 
#arrange(-n_studies_shared) %>% 
#mutate(Percent_total = n_studies_shared / n_distinct(data$Study_id))
#  
##Seelect quantile 50% from the total number of species
#spp_number = plant_spread %>%  
#select(Plant_accepted_name) %>% 
#n_distinct()
#df = seq(1, spp_number)
#quantile_50 = quantile(df, 0.5)
#
##Obtain most common plant for plotting
#most_common_plant = plant_spread %>% filter(n_studies_shared == max(n_studies_shared)) %>%
#select(Plant_accepted_name) %>% pull() 
#
#Plot
#p2 = ggplot(plant_spread, 
#aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
#geom_bar(stat = "identity", 
#           color = "#009E73",
#           lwd = 0.015, fill = "#009E73") +
#theme_minimal()+
#theme(axis.text.x = element_blank(),
#         axis.ticks.x=element_blank(),
#         axis.ticks.y = element_line(size = 0.2),
#         axis.title.x = element_text(face = "bold", size = 7),            plot.title = #element_text(size = 8, face = "bold"),
#         axis.text.y = element_text(size=5),
#         panel.grid.major.x = element_blank(),
#         panel.grid.minor.y = element_blank(),
#         panel.grid.major.y = element_line(size=0.2, linetype =  2),
#      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
#xlab("Species") +
#ylab("Occurrence across studies") + 
#ggtitle("Plants") + 
#geom_vline(xintercept = quantile_50, linetype="dashed", 
#          color = "black", size=0.25) + 
#annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
#annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
#      ,color="black",  fontface = 'italic', size = 1.5)+
#annotate(geom="text", x=20, y=-0.08, label="Common spp."
#      ,color="black",  size = 1.5) +
#annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
#      ,color="black",  size = 1.5) + 
#scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = #label_number(accuracy = 0.01)) +
#coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, #max(plant_spread$Percent_total)), clip="off")
#

data2 = data %>%  select(Study_id, Year) %>% 
distinct() %>% 
group_by(Year) %>% 
summarise(Counts = n())

p2  = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 7, vjust = -2),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
   axis.ticks.x = element_line(color="black"),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10,hjust = -0.3, vjust = -5))

main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")
#Select main orders
data1 = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Species = n_distinct(Pollinator_accepted_name)/n_distinct(.),
          Interactions = length(Pollinator_accepted_name)/ nrow(.))

# Combine Interactions and Species into a single column
data_long = pivot_longer(data1, cols = c(Interactions, Species), names_to = "Variable", values_to = "Value")

#Set order of levels for plotting
data_long$Pollinator_order = factor(data_long$Pollinator_order, levels = c("Coleoptera", "Lepidoptera", "Diptera", "Hymenoptera"))

>>>>

---
output:
  pdf_document:
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. ", fig.width=10, fig.height=7, fig.align='center'}
